### 2019-05-05 presentation/equations.tex
add equations

\documentclass[a4paper,12pt]{article}
\usepackage[fleqn]{amsmath}
\usepackage{amsmath}
\begin{document}
\noindent
{\setlength{\mathindent}{0cm}

\noindent \textbf{FOMC Decison $D_t$ based on lagged document text}
\begin{equation*}
E[D_t | \mbox{Statements}_{t-i}, \mbox{Minutes}_{t-j} , \mbox{Speeches}_{t-k}]
\end{equation*}

\noindent \textbf{FOMC Decison $D_t$ based on aggregated lagged document text}
\begin{equation*}
E[D_t | \mbox{AGG}(\mbox{Statements}_{t-i}, \mbox{Minutes}_{t-j} , \mbox{Speeches}_{t-k})]
\end{equation*}
\noindent where 
\begin{equation*}
D_t = 1, \mbox{if change to rates}
\end{equation*}
\begin{equation*}
D_t = 0, \mbox{if no chagne to rates}
\end{equation*}
\end{document}