### 2019-05-13 presentation/paperv2.tex
paperv2

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%
% ML Presentation Paper 
% Jeremy Lao/John Reynolds
% JJL359, JR4716
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\documentclass[11pt]{article}

% AMS packages:
\usepackage[fleqn]{amsmath}
\usepackage{amsmath, amsthm, amsfonts}

% Document Formatting Packages:
\usepackage{hyperref}
\usepackage[margin=1in]{geometry}
\usepackage{graphicx}
\usepackage{caption}
\usepackage{subcaption}
\usepackage{float}
\newcommand{\vertSpace}[1]{\vspace{3mm}}

%----------------------------------------------------------------
\title{Predicting FOMC Actions using NLP; Assesing the impact of matrix sparsity and regularization}
\author{
        Jeremy Lao - jjl359 \\
        NYU Computer Science \\
            \and
        John Reynolds - jr4716 \\
        NYU Computer Science \\
}

\begin{document}
{\setlength{\mathindent}{0cm}
\maketitle

\abstract{\textit{The Federal Open Market Committee (FOMC) meets throughout the year to set the target Federal Funds rate.  The Federal Funds rate is one of the primary monetary policy tools, and it impacts the cost of borrowing money globally.  This paper explores ML and NLP methods, examines the impact of matrix sparsity from NLP methods and regularization from ML models, to predict the potential outcome of an upcoming FOMC using past FOMC statements and meeting minutes to train ML and NLP models.}}

\section{Introduction}
In this paper we use methods in Natural Language Processing (NLP) and Machine Learning (ML) to predict Federal Open Market Committee (FOMC) rate actions (hold or change) using text from FOMC meeting minutes, Board of Governors speeches, and FOMC post-meeting statements.  
For this analysis we used 713 separate documents (rougly 1.6 Million Words) and 148 Fed Decisions for our data set.  
Employing NLP techniques can result in large numbers of features being generated and an extremly sparse resultant feature matrix.  
In the analysis we conducted some of the feature matricies generated contained over 2 Billion elements with fractions as small as 0.0025 of the elements containing a non-zero value.  
Our analysis and techniques used were driven by the issue of feature sizes orders of magnitued larger than the label size and the extreme degrees of sparsness in the feature matrix.  
To make sure the analysis was at least plausible we created a simulator to detect positive or negative sentiment from randomly generated text and found as the number of sentiment words increased the prediction accuracy approached close to 1.0.  
For the study the analysis hones in on the mix of distributions containing hawkish or dovish(positive) words.

\subsection{git repository for project}
\url{https://www.github.com/jrrpanix/ML9}

\section{Objective}
\begin{enumerate}
\item Since its pretty easy to just blindly apply tools to a set of data we wanted to understand the 
characteristics of the output from the text processing functions (in our case CountVectorizer) in terms of number of features, matrix sparsity, matrix norm
that are generated from calls to this function.
\item Assess the effeiveness of Bayesian Approaches and Regularization on the analysis.
\item Determine whether ML and NLP methods can be used to predict FOMC action (no change or a change in the base rate).
\item Because 70$\%$ of the actions are ``no action'', make sure the prediction is not always just ``no action'' (which would give a 70$\%$ accuracy).
\end{enumerate}

\section{Methodology}

\subsection{Data}

The time frame of our study covered 2000 to 2018. 
We primarily utilized three sources for our data, FOMC statements, FOMC meeting minutes and Boarde of Governors speeches.
Details on datasources, extraction and scrubbing are left to the appendix to keep the focus on the methods and results.
Preparing the data was a large amount of work and we are not aware of any data sets that combine the three types of documents
and then map the documents to a Fed Action.

\subsection{Feature Modeling and Data Label Mapping}
\noindent Documents are mapped to the first rate decision on or before the document date at time $ti$
where $j=\text{Statement}$ or $j=\text{Speech}$ or $j=\text{Minutes}$
\begin{equation*}
\mbox{Document}_{j,ti} \rightarrow \mbox{Decision}_{ti+k}
 = \begin{cases}
  1, & \text{if  Decision = change}, \\
  0, & \text{if  Decision = no change}
\end{cases}
\end{equation*}

\noindent 
\textbf{Model 1: Unstacked}\\
\noindent
The \textbf{Unstacked} model maps each document to a feature vector generated by CountVector given an n-gram($n_1$, $n_2$)
where $n_1 <= n_2$ and $n_i$ are integers $>0$. Given there are approximately 713 documents, this will generate
a feature matrix with 713 rows and $K$ columns where each column will be and instance of the word sequence
generated by the n-gram.

\begin{equation*}
E[D_ti] = f(ngram(n_1, n_2)[\mbox{Document}_{j,ti-k}])
\end{equation*}

\noindent
\textbf{Model 2: Stacked}\\
The \textbf{Stacked}  model maps the set of documents which occurr after the previous decision date but on or before the
next decision date to one rate decision.  This has the effect of contatenating the documents between decisions into
one large document, but cuts down the number from 713 to 148.
where $n_1 <= n_2$ and $n_i$ are integers $>0$. Given there are approximately 713 documents, this will generate
a feature matrix with 713 rows and $K$ columns where each column will be and instance of the word sequence
generated by the n-gram.

\begin{equation*}
E[D_{ti}] = f(ngram(n_1, n_2)[\mbox{Document}_{j1,ti-k1},{Document}_{j2,ti-k2},...]) \text{, } ti-1 < ti-ki < ti+1
\end{equation*}

\subsection{Matrix Sizes and Sparsity from CountVector}
\noindent For this study we used \textbf{CountVector} (\textbf{TfiDf} can also be used, but it did not alter the conclustions of this study).
When are n-gram is applied to corpus it combinatorics diticates a large number of features will be generated and for a given feature, most of the column entries will
be zero, resulting in extremly sparse matricies.  The table below gives a measure of $\mbox{sparsity} = \frac{non-zero elements}{total elements}$ and 
the resulting number of elements in each n-gram generated matrix in the billions. In general stacking reduces the sparsity by a factor of 4 and
the matrix size by a factor of 5.\\

\noindent \begin{tabular}{ |p{2cm}||p{2cm}|p{2cm}|p{2cm}|p{2cm}|  }
 \hline
 \multicolumn{5}{|c|}{CountVector Matrix Size} \\
 \hline
 \multicolumn{1}{|c|}{} &
 \multicolumn{2}{|c|}{Unstacked} &
 \multicolumn{2}{|c|}{Stacked}\\
 \hline
 N-gram & sparsity & size(billions) & sparsity & size(billions)\\
 \hline
       1:1& 0.03628801& 0.008561&  0.10635940&   0.001757 \\
       1:2& 0.00544312& 0.170058&  0.02081416&    0.036107\\
       1:3& 0.00368457& 0.439898&  0.01508941&     0.092869\\
       3:5& 0.00238128& 0.900721&  0.01059293&     0.173040\\
       4:7& 0.00224699& 1.332560&  0.00996789&     0.271117\\
      5:10& 0.00218455& 1.995502&  0.00972276&     0.402041\\
      8:10& 0.00213633& 1.019811&  0.00955894&     0.198627\\
     10:15& 0.00210643& 2.097187&  0.00945586&     0.398312\\
     20:20& 0.00205055& 0.347043&  0.00931157&     0.074799\\
 \hline
\end{tabular}

\section{Conclusion}

We found that the model results and methodology for simulating training and test data for model and feature selection to be robust.  Our next recommendation is to utilize this in the production environment and evaluate whether there are trading opportunities when predicting the probability of a move versus the market implied probability from Fed Funds futures.

\appendix
\section{Appendix}

\subsection{What is the FOMC}

The FOMC sets the Federal Funds target range (target rate prior to 2009).  The permanent members are the Board of Governors of the Federal Reserve System, President of the Federal Reserve Bank of New York, and the rest of the seats are rotated through the Presidents of the other Federal Reserve Banks.  There are twelve Reserve Banks and the Board of Governors oversees the activity, operations, and policies of the Federal Reserve System.

\subsection{Why Does the Federal Funds Target Range Matter?}

The Federal Funds Target Range is set by the FOMC.  The Federal Funds Target Range is the range where the Federal Funds Effective rate (calculated as the volumn weighted median of eligible transactions reported on FR 2420) is exepected to fix on a daily basis.  The Federal Funds rate is the primary monetary policy instrument of the Federal Reserve System, and it influces the level of interest rates domestically and globally.  For example, if the Federal Funds effective rate were 5 percent, then the interest rate on a 30 year mortgage would be 5 percent or greater.

\subsection{Data}

\subsubsection{Data Sources}
The time frame of our study covered 2000 to 2018.  
The primary source of our textual data were FOMC statements, FOMC meeting minutes, and Board of Governors speeches from \url{https://www.federalreserve.gov}.  
For FOMC statements between 2000 and 2008 we utilized Stanford's FOMC corpus.  
The Stanford FOMC corpus only has text prior to 2008 and it does not have Board of Governor speeches.
The data for the Federal Funds target rate and target range (post 2009) is from FRED St. Louis.  
FRED offers a wealth of economic data and information to promote economic education and enhance economic research. 
FRED is updated regularly and allows 24/7 access to regional and national financial and economic data.

\subsubsection{Scraping and Text Pre-Processing}

We collected data from the Federal Reserve website and pre-processed the data for the document-term matrix. \vertSpace

We utilized Python packages \textit{beautifulsoup4}, \textit{re}, and \textit{urllib} to scrape contextual data from the Federal Reserve website.  Our scraping algorithm used regular expressions to handle and remove extraneous \textit{html} and \textit{javascript} text collected by the page scraper.  We had to handle non UTF-8 characters upfront and remove them at this stage altogether.  We collected over six hundred documents from the Federal Reserve website. \vertSpace

After scraping the data, we pre-processed the text by removing all punctuation, ensuring proper spacing between words, setting all words to lower case, and making all numbers \textit{d} to reduce the dimensionality document-term matrix.  We also used Regex to detect direct references to the Federal Funds target rate/range and transformed those references (a mixture of numbers, punctuation, and the word 'percent') to \textit{ddpercentrate}.  Since we were modeling the sentiment of FOMC rate action, this was one of our strategies to directly capture the target rate in the document-term matrix. \vertSpace

This formed the bulk of our text pre-processing and it allowed us to essentially utilize the stored text in the document-term matrix for NLP and ML models.

\subsubsection{Meta-Data}
\noindent Below is a table of the FOMC document metadata:

\vertSpace

\noindent \begin{tabular}{ |p{2cm}||p{2cm}|p{2cm}|p{2cm}|p{2cm}|p{2cm}|  }
 \hline
 \multicolumn{6}{|c|}{FOMC Documents} \\
 \hline
 Document Type& Years & Num Documents & Total Words & Avg Words Per Document & StDev\\
 \hline
 Minutes   & 2000-2018    & 149 & 738,656 & 4,957 & 1,284\\
 Statements &   2000-2018  & 160   & 62,350 & 389 & 185\\
 Speeches & 2011-2018 & 403 & 915,359 & 2,271 & 1,539\\
 \hline
\end{tabular}

\vertSpace

Statements are post-meeting communiqu\'e and are generally short, but contains a summary of the committee's reasoning for the decision.  Minutes are longer and contain the Staff and Committee member current economic assessments and outlooks.  The speeches in the corpus are from the Board of Governors and may contain clues of their thoughts on their current economic outlook.  \vertSpace

\subsection{Latent Dirichlet Allocation}

We explored latent topic modelling in our research to determine how a machine learning algorithm would classify $k=3$ topics, a proxy for the Federal Reserve's three objectives. \vertSpace

Given the Federal Reserve's mandates of stable prices, maximum employment, and financial stability, we explored Latent Dirichlet Distribution of the speeches and minutes' word distribution over three latent topics.  We were interested in exploring how well a topic model would distribute words over the three objectives and if the meaning of the distributions were relatively clear.

\begin{figure}[h!]
  \includegraphics[width=\linewidth]{../output/data_for_graphs/Topic-Model-3-topics.PNG}
  \caption{Top 15 Words Distributed Over Three Latent Topics}
  \label{fig:LDA}
\end{figure}

We found that the topic model generally distributed the speeches and minutes into the three categories fairly well.  It was interesting that "Topic 1" referred solely to financial stability, "Topic 0" primarily referred to inflation and spending, and "Topic 2" returned a word distribution that included labor, growth, and lower for longer monetary policy. \vertSpace

% Bibliography
%-----------------------------------------------------------------
\begin{thebibliography}{99}

\bibitem{Cd94} Hansen,Stephen, McMachon, Michael  \emph{Transparency and Deliberation with the FOMC:a Computational Linquistics Approach}, CEPR Discussion Paper, (2014)

\bibitem{stanford} Puri, Indira \emph{Using Machine Learning to Predict Interest Rate Changes from Federal Reserve Proceedings}

\bibitem{nyu} Ranganath, Rajesh  \emph{NYU - CSCI-GA.2565 ML Lecture Slides Spring 2019}
\end{thebibliography}

\end{document}